### 2013-11-21 paper/paper.tex
mostly rewrite pill section, and start synchronizing box plot with image plot Tested everything on collins.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}
\usepackage{graphicx}% Include figure files
\usepackage{color}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}

\newcommand{\fixme}[1]{\red{[#1]}}
\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\renesays}[1]{{\color{red} [\blue{Rene:} \emph{#1}]}}
\newcommand{\jeffsays}[1]{{\color{red} [\blue{Jeff:} \emph{#1}]}}

\newcommand\micron{\ensuremath{\mu\text{m}}}

\begin{document}
\title{Robustness of MinD oscillation in \emph{Escherichia coli} with
  diverse cell shapes}

\author{Jeff B. Schulte}
\affiliation{Department of Physics, Oregon State University}
\author{Rene W. Zeto}
\affiliation{Department of Physics, Oregon State University}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University}

\begin{abstract}
  The dynamics of the Min-protein system help Escherichia coli
  bacteria regulate the process of cell division by identifying the
  center of the cell.  We model the Min-protein system in bacteria
  that have been forced into unusual flattened shapes, as have
  recently been experimentally observed.  We find that although the
  presence of Min oscillations is robust in a wide variety of cellular
  configurations, the location of the peaks is strongly affected by
  the cellular shape.  In some cases no periodic oscillations are
  observed.  In particular, we find that cellular shapes observed
  experimentally to present irregular oscillations do so in the
  theoretical model, consistently \fixme{or inconsistently?} with
  experiment.  \fixme{In agreement with previous theoretical and
    experimental results, we observe ``rotating'' behavior in certain
    shapes having three corners.}
\end{abstract}

\maketitle

\section{Introduction}
It is vital that during the process of bacterial cell division, a cell
avoid minicelling, or splitting into daughter cells with lopsided
volumes.  A long FtsZ polymer chain develops on the cell wall, in the
center region of the cell, the region which dictates the plane of
division~\cite{adams2009bacterial, lutkenhaus2007assembly}. Previous
experimental studies have shown that the MinC protein, known to
inhibit the FtZ polymer\cite{shen2010examination}, exhibits pole to
pole oscillatory behavior in conjunction with the MinD and MinE
proteins, while the oscillating MinE tends to be more localized in the
center of the cell~\cite{hu1999topological, fu2001mine,
  shapiro2009and, yu1999ftsz, raskin1999rapid}. The MinC protein will
then have a higher time averaged concentration in the cell poles, as
opposed to the center region of the cell, aiding in prohibiting the
FtZ from developing in the wrong region.

The cell division process must be able to prevent minicelling, even
under irregular conditions. Studies have shown that it is capable of
performing this task for fairly perturbed cell
shapes\cite{touhami2006temperature}. Mannik et al. have done some
early experimental work showing regular oscillations of cell division
systems in bacteria that were physically forced into unusual
shapes. Varma et al. studied one particular example of a perturbed
shape, a three-pronged tube, and found both experimental and
simulation results\cite{varma2008min} supporting the idea that certain
systems inside the cell begin to undergo oscillations inside the
cell. These oscillations seem to seek out the extreme poles in the
cell shape\cite{corbin2002exploring} \cite{juarez2010changes}, and
effectively prevent minicelling in most cases. However, Mannik et
al. have also shown that there are limitations to this
mechanism\cite{mannik2010bacteria} \cite{mannik2009bacterial}. One
instance of this behavior is observable when the cell has been forced
into a flattened, irregular shape. Oscillations in this sort of cell
shape are far from the regular sorts of oscillations seen in the
computational work done by Huang et al\cite{huang2003dynamic}. We seek
to apply a computational model to explore the more extreme or
irregular cell shapes, in order to determine the limits of the steady
cell oscillations seen by Mannik et al.

%is this necessary? was in old paragraph.
%\fixme{this link doesn't work =
%  cite{mannik2012robustness}} \cite{mannik2010bacteria}
%\cite{mannik2009bacterial}

A significant amount of work has been done to develop protein reaction
and diffusion models that exhibit accurate macroscopic dynamics of the
MinD protein system. Early models involved free proteins that affect
each others' rates of diffusion and membrane attachment, but do not
combine into compound states~\cite{meinhardt2001pattern}.  In 2003
Huang improved upon this work with a simple model based on MinD-MinE
combination, ATPase hydrolysis, and MinD membrane attachment that when
simulated, exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is more
likely to attach to the membrane when MinD is already clustered there
(following observed non-linear attachement of minD on the cell
membrane), and is stationary once attached.  A number of studies have
used an approach similar in that they do not rely on the ability of
MinD to move along the walls and
cluster~\cite{kruse2007experimentalist, meinhardt2001pattern,
  drew2005polymerization, fange2006noise, kerr2006division}, while
studies have been made as well of models which rely on MinD mobility
and attraction on the cell membrane~\cite{kruse2002dynamic,
  howard2005cellular}.  Variations of the Huang 2003 model that
stochasticly account for variations of molecular interaction
\cite{fange2006noise} and monte-carlo simulations that implement this
stochastic version of Huang's mean field reaction rates confirm the
major results obtained by Huang's model, and more successfully predict
experimentally observed oscillations in round cell
phenotypes~\cite{drew2005polymerization, fange2006noise,
  huang2004min}. Biochemical models of broader scope have also been
used to study the MinD system and show consistent
results.\cite{arjunan2010new}.  However, in general, the results of the
stochastic and monte-carlo simulations are similar to those given by
Huang's mean field results.

Previous studies have been made of the Min system's associtation with
the cell membrane
\cite{hsieh2010direct}\cite{mileykovskaya2003effects}.  Studies have
shown as well that MinD binds preferentially to regions enriched with
cadiolipin, an anionic phospholipid that collects on regions of high
negative curvature. This mechanism has been incorporated into other
models.\cite{drew2005polymerization,cytrynbaum2007multistranded,renner2012mind,renner2012mind}
However, this mechanism of combined clustering, phospholipids and MinD
has not been observed in real cells. \cite{halatek2012highly}

\begin{figure}
  \includegraphics[width=\columnwidth]{reactions}
  \caption{Reactions included in the model of Huang \emph{et
      al.}~\cite{huang2003dynamic}.}\label{fig:reactions}
\end{figure}

\section{Model and Methods}
We implement the model of Huang \emph{et al.}  for the behavior of the
MinD and MinE proteins inside the cell~\cite{huang2003dynamic}.
Figure~\ref{fig:reactions} shows the reaction process.  The
cytoplasmic MinD:ADP complex undergoes nucleotide exchange and is
changed into the MinD:ATP complex.  This will naturally diffuse
towards and attach to the cell membrane.  A cytoplasmic MinE will
attach to the wall bound MinD:ATP complex and after a time will
activate ATP hydrolosis.  This breaks up the complex, releasing MinE,
phosphate, and MinD:ADP back into the cell.  The MinD:ADP will undergo
nucleotide exchange and begin again the cyclic process.  This model is
defined by a set of five reaction-diffusion equations:

\begin{multline}
  \frac{\partial \rho_{D:ADP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ADP}-k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  +\delta(d_w)k_{de}\sigma_{DE},\hspace{3.4cm}
\end{multline}
\begin{multline}
  \frac{\partial \rho_{D:ATP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ATP}+k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  -\delta(d_w)[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
\end{multline}
\begin{multline}
  \frac{\partial \rho_E}{\partial t} = \mathcal{D}_E\nabla^2\rho_E+\delta(d_w)k_{de}\sigma_{DE}
  -\delta(d_w)k_E \sigma_D \rho_E
\end{multline}
\begin{multline}
  \frac{\partial \sigma_d}{\partial t} = -k_E\rho_d\rho_E
  +[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
  \label{eq:d-on-wall}
\end{multline}
\begin{multline}
  \frac{\partial \sigma_{DE}}{\partial t} = -k_{de}\sigma_{DE}+k_E\sigma_D\rho_E\hspace{3cm}
  \label{eq:FifthPDE}
\end{multline}

where $\rho$ is cytoplasmic protein density ($\mu m^{-3}$), $\sigma$
is membrane bound density ($\mu m^{-2}$), $\mathcal{D}_D$ and
$\mathcal{D}_{E}$ th diffusions constants for MinD and MinE,
respectively, $k_D^{\textrm{ADP $\rightarrow$ ATP}}$ the rate of
conversion from MinD:ADP to the MinD:ATP complex, $k_D$ the rate of
MinD:ATP attachement to the membrane when no protein is already
attached there, $k_{dD}$ the increase of this rate when MinD:ATP is
present on the membrane, $k_{de}$ the rate of hydrolisis of the
MinD:MinE:ATP complex, $k_E$ the rate of cytoplasmic MinE binding to
membrane bound MinD:ATP complex, and $d_w$ is the distance from the
point in space to the closest wall (which will always be
perpendicular to this distance).  $\delta(d_w)$ has units of
$\mu m^{-1}$ and will be zero everywhere except at the wall.
Equations \ref{eq:d-on-wall} and \ref{eq:FifthPDE} will only be
relevent at the membrane because the membrane bound density values
will have no meaning in the cytoplasm where there is no membrane.

Our diffusion and reaction rates are shown below.  We are interested
primarily in the effect of cellular shape on the protein
oscillations, so we follow Huang's work\cite{huang2003dynamic} and do
not deviate from the wild type values used in that work.  Huang's
simulations use total MinD and MinE concentrations of $1,000/\mu m$
and $350/\mu m$, respectively, in a cylindrical cell of radius $0.5\mu
m$, and in our (non-cylindrical) cells we use the same number of
proteins per unit volume.  These concentration values are 1273 $\mu
m^{-3}$ and 445 $\mu m^{-3}$, respectively.
\begin{gather*}
  \mathcal{D}_D = \mathcal{D}_{E} = 2.5\micron^2/\text{sec}\\
  k_D^{\textrm{ADP $\rightarrow$ ATP}} = 1/\textrm{sec,  }
  k_D = 0.025 \micron /\textrm{sec}\\
  k_{dD} = 0.0015 \micron^3/ \textrm{sec,  }
  k_{de} = 0.7/\textrm{sec}\\
  k_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}
A 3D grid is then constructed in cartesian coordinates, with a grid
spacing of .05\micron. We define our cell shapes and solve the
reaction-diffusion partial differential equations numerically to
observe the time evolution of the MinD and MinE concentrations inside
the cell.

The reactions shown in Figure~\ref{fig:reactions} are cyclical so we
can begin our observations at any stage in the cycle (which is
presented for the natural, pill shapes cell in Fig.~\ref{image-p}
below).  We consider first MinD:ATP floating freely in the cytoplasm.
The protein will not react with any other floating protein, and it
will not spontenously shift back into the MinD:ADP stage of the cycle.
It will diffuse until it attaches to a cell membrane.  This will
happen most rapidly in cytoplasmic regions that are close to membrane
regions of high curvature, where diffusion in a random direction is
more likely to lead to an encounter with a wall.  This effect is made
more pronounced by the MinD:ATP tendency to attract and cluster on the
membrane, as reflected in Equation~\ref{eq:d-on-wall}, which shows
that when adjacent to the membrane, the rate of binding to the
membrane is porportional to the amount of MinD already bound to the
wall.  Initially, then, there is a strong tendency of MinD to cluster
on the polar regions of the cell, where the curvature is highest.

We next see the formation of MinE `rings' on the membranes of the
polar caps.  As the freely diffusing MinE enters the cap portion it
begins to encounter concentrated membrane bound MinD.  The MinE binds
to this MinD and remains for a time on the walls in compound form.
When it is released, it has the two options of diffusing naturally
outward or inward towards the cell.  If it diffuses inward, there is
less cross sectional wall for it to bump into and also less
concentrated MinD:ATP to react with.  If it diffuses outward, the
curvature of the cell wall cuases there to be a greater probability
per unit time that it will encounter the cell membrane (for the same
reason outlined for the MinD:ATP above) where there is more membrane
bound MinD:ATP. If it does diffuse in this directin, it will likely
react with MinD:ATP before getting very far into the encap.  This
process results in the observed qualitative 'encroaching MinE ring'
behavoir that releases MinD:ADP back into the cytoplasm.  This
MinD:ADP will diffuse for a time before spontanosly undergowing
nucleotide exchange and reaching the MinD:ATP state that
can bind to the membrane again.  There is a strong chance that it has
by this time diffused out of the polar region and is on its way to the
opposite polar region of the cell.  If it hasn't then it will reattach
to the MinD:ATP on the cell wall deeper into the endcap and will
eventually be converged upon by the MinE ring, only to be released
again in MinD:ADP form.

Overall this process results in extremely robust oscillatory
behavoir of polar selection and oscillation in a wide variety of cell
shapes.

\section{Cell shapes}

We mean to investigate the geometric limits of the Min system
oscillations, as observed by Mannik \emph{et
  al.}~\cite{mannik2012robustness}, so we model the Min system in a
variety of cell shapes.  We begin with naturally occuring pill-shaped
cells and then proceed to examine a number of flattened pancake-like
shapes, which reflect the experiment of Mannik \emph{et al.}, in which
bacteria were confined within a thin slit~\cite{mannik2012robustness}.
Our pill shapes differ from those of Huang \emph{et al.} in that they
are cylinders with hemisphere endcaps instead of pure cylindrical
shapes.  Our cylinderidrical radius is $0.5\micron$ and the lengths
of our cells (measured between the tips of the endcaps) are
$5\micron$, $4\micron$, $3\micron$, and $2.5\micron$.

With our flattened cell shapes we mean to study cells that are
parrallel to those lodged into crevices by Mannik so we give them
height of $0.25\micron$ \cite{mannik2012robustness}.  Viewed from the top
down the cells will have the shapes described below and viewed from
the side they have at their edges a semicircular protrusion (one may
imagine the edges of a pancake).

Our first group of (three) flattened cell shapes are triangular when
viewed from above.  One of them is equilateral, one has a 3-4-5 right
angle shape, and the last is iscoscolese.

Our second group of (two) flattened cells are designed to replicate
those published by Mannik, in order to compare the results of the
simulation model with the experimental results in that work.

Our third group of (three) flattened cells are our own creations,
designed to investigate different patterns of protein oscillation
behavoir.

Kubitschek has shown in multiple experiments that at the time of cell
division cells have a volume that is within a range of roughly
$1\micron^3$ to $2\micron^3$~\cite{kubitschek1990cell,
  kubitschek1968linear}.  We follow Huang's
simulations\cite{huang2003dynamic} and Mannik's experiments and model
cells that are slightly larger than this range.  Amoung our flattened
cells the two dimensional length scales are tuned so that total cell
volumes are very close to $3\micron^3$. \jeffsays{I'm not sure how
  true this is.  We should go through and check.  There've been a few
  implementations of printouts reporting volume in the cpp}
\section{Specific Results}

\subsection{Pill Shape}

\begin{figure*}
  \includegraphics[width=\textwidth]{../data/shape-p/plots/image-plot--p-400-50-0-0-1500.pdf}
  \caption{Images of the concentration of each species in a natural
    pill-shaped bacterium as a function of time, with darker regions
    indicating higher concentration.  Frames from an animation showing
    a series of contour plots of the concentration of proteins in
    different stages over time.  The order of frames is such that
    individual MinD proteins begin at the bottom of the plot (in the
    MinD:ATP state in the cytoplasm), and progress upward until they
    reach the MinE:MinD:ATP membrane-bound complex.  At that point,
    they will spontaneously dissociate into cytoplasmic MinE (the top
    row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.}
  \label{image-p}
\end{figure*}

We begin with the naturally occuring pill cell shape.  We peice this
shape together as two hemispherical endcaps attached on either end of
a cylinder.  This shape follows the early simulations of Huang
\emph{et al.} but differs in that we have added the end caps for a
more natural shape, expecting similar results.  We test cells of
radius $.5\micron$ and of lengths $2\micron$, $3\micron$, and
$5\micron$ measured from the tip of one end cap to the top of the
other. In each case, we observe quick establishement (within a few
periods) of extremely regular oscillatory behavoir that lasts
indefinitely.  The period increases with cell length, as shown in
Table~\ref{tab:pill-periods}, in agreement with the results of Huang
\emph{et al.}.

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline
    Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\
    \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2}
    \hline
  \end{tabular}
  \caption{Period of oscillation according to length of cell for
    pill-shaped cells.}\label{tab:pill-periods}
\end{table}

This simple model is a good starting point for observing in detail the
dynamic interaction between the different stages of protein that lead
to their qualitatively distinct behavoir. Figure \ref{image-p} shows
the process in a series of frames taken from simulation animations.
Each frame is 2.5 seconds ahead of the last, and each image shows a
the concentration of a given state of protein summed over the
coordinate orthogonal to the page.  The color scale for each protein
state is set according to the maximum density observed.

In the beginning, the cytoplasmic MinD:ATP is roughly evenly
distributed throughout the cell with close to its peak density.  There
is a high concentration of MinD on the membrane (in both its
membrane-bound states) at the lower end of the cell, and there is also
cytoplasmic MinD:ADP concentrated in the lower end, which has recently
left the cell membrane.   There is relatively little cytoplasimic
MinE, as it is primarily bound to the lower end of the membrane.

After ??? seconds, very little MinD remains bound to the membrane, and
there is a high concentration of cytoplasmic MinE in the lower half of
the cell.  In the mean time MinD has begun binding in significant
quantities to the upper half of the cell membrane, resulting in a
diminishing of the concentration of cytoplasmic MinD:ATP in the upper
half of the cell.  At this stage, we begin to see the formation of the
``MinE ring'' in the center of the cell---visible here in the plot of
the concentration of membrane-bound MinE:MinD:ATP as two lines
representing a cross section of the ring.  This ring forms as the
cytoplasmic MinE reacts with the membrane-bound MinD:ATP, and leads to
the removal of MinD from the membrane.

From ??? seconds to ??? seconds, we see the MinE ring progress
upwards, as the cytoplasmic MinE diffuses in that direction,
progressively removing the membrane-bound MinD from eventually all but
the very top of the cell membrane.  During this process, the MinD:ATP
bound to the top of the cell membrane is attracting increasing
quantities of MinD:ATP from the cytoplasm.  At ??? seconds, the cell
is in the mirror image state from which it started, and the period is
half-complete.  The second half of the period is a mirror image of the
process just observed, as MinD is driven from the membrane at the top
of the cell, begins binding to the lower half of the cell and a MinE
ring pushes it back down to lower end of the cell.

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-400-50-0-0-1500}
  %\includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-200-50-0-0-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of a 5$\micron$ by 1$\micron$ pill shaped cell (above) and a
    3$\micron$ by 1$\micron$ pill shaped cell (below).  The vertical axis shows
    stacked the total number of proteins that are of four different
    compound stages and in the right section of the cell (bottom four
    colors), in the middle of the cell (middle four colors) and in the
    left section of the cell (top four colors). The different forms of
    protein changes while the total number of MinD protein in the cell
    remains constant, shown by the fact that the very top line is
    horizontal.}
  \label{box-p}
\end{figure}

\davidsays{We should make Fig. \ref{box-p} and Fig. \ref{image-p} plot
  for precisely the same times, that will make it way easier to read
  the two plots.}  While Fig.~\ref{image-p} is helpful for visualizing
the process occuring in the Min system, it is does not show the
relative number of proteins in each state.  To visualize this
information, we plot in Fig.~\ref{box-p} the fraction of MinD found in
each stage of the cycle in each of three sections of the cell---top,
middle and bottom---as a function of time over one period.  The time
axis is the same as in Fig.~\ref{image-p}.  Looking at the evolution
of the latter, we see an initial high concentration of cytoplasmic
MinD:ATP, as the this protein diffuses into the polar region, followed
by a dip in this and large spike in membrane bound MinD:ATP, as the
protein clusters onto the polar cell wall, followed by a spike in the
memebrane bound MinE:MinD:ATP stage of protein, together with a slight
rise in cytoplasmic MinD:ADP as the MinD dettaches from the wall.

results for pill shaped cell of length
$5\mu$ and radius $.5\mu$.  The plot shows both the transerence
between different stages of protein compound and movement within the
cell.  Starting at the beginning of a period, we can see that first
there is a spike on the left of MinD-ATP attached to the wall that is
accompanied by a growing number of cytoplasmic MinD-ATP and MinD-ADP
that are entering the left section, followed by a high peak in
MinD-MinE-ATP.  This peak gives way through the middle to peaks on the
right side of the cell, the first being a small bump in cytoplasmic
MinD-ATP that dissapears into a much larger spike in wall attached
MinD-ATP, which is followed by a peak in wall attached MinD-MinE-ATP.
The pattern during each period clearly shows the cycle of proteins
rections, as well as the very regular nature of the pole to pole
protein oscillations.

\fixme{our 5 length cells are labeled as 4 cells in the file labels,
  since 4 is the length of the cylinder. In the text I'll refer to the
  5 length cells (file says 4) as 5 length cells.  What huang's paper
  calls 4 length cells will be labeled by our files as 3 length cells,
  but will be refered in our text as 4 length cells}.  We can see a
clear difference in period between the $5\mu$, $4\mu$, and $2\mu$
length cells.  The oscillation periods for our $4\mu$ show excellent
agreement with the simulations of Huang et all
\cite{huang2003dynamic}.  This is to be expected since we begin our
simulations with their reported wild type concentrations of MinD and
MinE proteins. The periods for the different pill lengths are shown in
the table below.  We can see a direct correltation between size and
period, as the proteins must diffuse further in a longer cell before
they accumulate on the opposite wall.

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/ave-plot_D--p-200-50-0-0-1500}
  \caption{Average protein density per unit area of wall for the wall
    attached stages of the protein cycle.  Shown in the right, middle,
    and left parts of a 5$\micron$ by 1$\micron$ pill shaped cell (above)
    and a 3$\micron$ by 1$\micron$ pill shaped cell (below).}
  \label{ave-per-area-plot-pill}
\end{figure}

Figure \ref{ave-per-area-plot-pill} shows the ave protein per unit
area that is wall attached in the middle, right, and left sections of
the pill shaped cell.  This is an important test of the central idea
behind the utility of the Min protein system.  The system is meant to
place Min proteins on the walls in heavier concentrations in the caps
of the cell and lower concentrations in the middle section, so that
the FtZ polymer may build up along the center walls.  We can see from
the plot of the 3$\micron$ by 1$\micron$ cell that the MinD proteins reach
concentration peaks in the caps that are higher than peaks in the
center.  \fixme{is the next part true?} Integrals of these plotted values show that the time averaged
concentrations are likewise higher in the caps than in the center.

\subsection{Triangular Shapes}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-400-400-400-1500}
  \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-500-300-500-1500}
  \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-600-480-360-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of an equilateral (top), iscoscolese (middle), and 3-4-5
    (bottom) triangle.  The vertical axis shows stacked the total
    number of proteins that are of four different compound stages and
    in the different sections of the cell.}
  \label{box-triangle}
\end{figure}

The double pole oscillation pattern, in which one pole is located in a
single corner of the triangle and the other pole is spread out between
the opposite two, can be seen in Figure \ref{box-triangle}.  The
equilateral triangle shows a spike in both the Center and Left
sections at the same time, both showing the familiar sequence of peaks
amoungst the different states of the protein seen in the poles of the
pill shaped bacteria.  A spike in the other corner, showing the common
sequence, occurs during the minima of the former two, wth a higher
concentration at the peak.  A similar pattern is seen in the
iscosolese triangular shape, where interestingly the onset of the
maxima in the single maxima corner (the Left section) builds more
quickly than it fades.  This is presumably becuase as the cytoplasmic
minD:ATP diffuses down a more accute corner, it is sourrounded by
closer cell walls than it would be in a less acute corner, so that the
it is more apt at any diffusive step to make contact and cluster on
the walls.  That particular part of the process is therefore
accelerated, driving the build up of the maxima forward.

The right triangle shows a similar pattern except that here there is
an assymetry between the two corners that maximize at roughly the same
time.  One corner is further from the single maximizing corner and it
consisitently peaks later.  The MinD simply have further to travel
before they maximize there.

For these three sizess and shapes there is not sign that the
oscillation patterns once they are set will break down.

\begin{figure}
%%  \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-600-600-600-1500}
  \caption{Total protein fluctuation of an equilateral triangle with
    larger dimensions (each side is roughly $6\micron$.  The vertical
    axis shows stacked the total number of proteins that are of four
    different compound stages and in the different sections of the
    cell. The time limits have been set to show the broken down
    section of the simulation.}
  \label{box-triangle}

The larger equilateral shape shows the same pattern as the smaller
version does only until fifteen periods, at which point there is a
breakdown of the oscillation pattern.  During the last three stable
periods, the single maximizing corner, the lower left section in the
figure, has a lower peak each period as the membrane bound MinD:ATP in
the other corners lag more and more before being eaten away by the
MinE.  Finally these maxima lag so much that they break down the
regular pattern.  There then occurs about five periods of semi chaotic
behavoir, until it settles back into another stable oscillation that
is similar to the original one, except that a different corner has
been chosen to be the singlular maxima (the right section in the
figure), while the other two corners maximize at the same time.  This
interesting breakdown and recovery shows both the robustness of the
double pole oscilatory tendency and the importance of required
distance of travel for the different proteins.

\end{figure}

Fifteen

\subsection{Randst shape}
\fixme{Cheat sheet for the shapes, need to put these into a big figure
  from the sections output: 96=star shape, 97=sideways sideways
  flying saucer pancake, 98=manniks squished cell, 99=an A}

\subsubsection{Randst-mannik shape}
\begin{figure}
  \includegraphics[width=2.8cm]{../mannik-2.png}
  \includegraphics[width=2.8cm]{../mannik-1.png}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1300-2000-9400-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1101-1101-9500-1500.pdf}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-600-800-9800-1500}
  \includegraphics[width=3cm]{../data/shape-randst/plots/arrow-plot-NflD-randst-25-1102-1102-9500-1500}
  \caption{Arrow plots for our simulation and mannik's experimental}
  \label{arrow-compare-mannik}
\end{figure}

Figure \ref{arrow-compare-mannik} shows a comparison of time maxima
plots with similar plots published by mannik et al.  The sequence of
arrows show a series of global maxima in space and local maxima in
time at their heads, numbered by the sequence in which they occur.
They function as a trace of the path that the traveling maxima take.
We have created for each of the two shapes published by Mannik a
closely resembling shape (shown directly below Mannik's) and a
variation on this shape (shown below that) in order to test the effect
of variations on basic shapes.  The plots of the left shape show that
much of the preffered locations for the proteins to maximize are
similar between this experimental and simulated shape.  The maxima in
the top center of the cell does not seem to appear in the experimental
version.  Also in both of the simulations on the left there don't seem
to be maxima at membrane regions with negative radii of curvature
while the experimental shows a maxima in the bottom center region
right at one of these regions.  The shape on the right compares less
successfully.  Here the simulation shows a limited number of maxima
locations that appear to be quite regular, where mannik's experimental
results show a much more chaotic system.  In particular, the maxima
location on the right wall seems to be very robust.  It acts as one of
the poles do in the pill shapes.  \fixme{I'm now running a sim with a
  slightly different shape to see result} The simulated version has
found a robust two pole pattern, similar to that seen in other shapes.

\subsubsection{Randst-99 our lambda shape}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9900-1500}
  %%\includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9600-1500}arrow-plot-NflD-randst-25-600-800-9800-1500.pdf

  \caption{Total protein fluctuation in the A shaped shape.  The vertical axis shows stacked the total number
    of proteins that are of four different compound stages and in the
    different sections of the cell.}
  \label{total-oscillation-randst-96-plot}
\end{figure}

Our animations for this cell shape show a very interesting pattern.
The maxima oscillates from the right corner to the top corner, then
\emph{back to the right corner} then to the top corner again, then to
the left corner, then back to the top, then to the left, then to the
top again, then back to the right side.

\fixme{running larger and smaller.  Guessing the larger will do pole
  and pole, so talk about in terms of you have huge, pole and pole,
  then once get small enough double oscillations develop.  Why do they
  do that?  And hen you can do image plot for the one with consistent
  double oscillations.  Can right now explain reason for double
  oscillations.  Look at todo.txt.  The slightly varied guassian shape,
  so a little assymetric, doesn't change double oscillation pattern.}

Contour plot frames from the animations show the second maxima in the
lower left corner (immediately following the one in the upper middle
corner) will reside for a bit longer than the left corner maxima
preceding it.  These longer residing maxima will then be followed by a
switch, to the right corner, with a short, less intensive maxima
traveling through the center.  The MinE shows the same pattern. The
MinE that was just concentrated on the left corner during the first
maxima, after releasing the MinD from the membrane there, diffuses up
and attaches to the MinD on the upper corner membrane and begins to
release it.  During this process the MinE starts at the edges and
moves in towards the center of the maxima.  The center of the cup is a
little further away so it attaches to these outer ridges first and
then moves in.  After it eats at the corners and starts moving in the
MinD eventually travels back to set up a maxima in the left corner.
Watching the cytoplasmic MinE, something very interesting: It as well
maximizes twice in the left corner, but during the MinD's first maxima
in the left corner, the MinD has a maxima in the upper corner, and
during the MinD's second left maxima, the MinE travels to the right
corner and has a maxima there.  At the same time (during the second
MinD left corner maxima), there is a large amount of cytoplasmic
MinD:ATP on the right side.  The MinE does this becuase during the
first MinD left corner maxima, the MinE has just come off a maxima in
the right corner.  So it diffuses up to the upper corner and sets up
there.  During the second MinD left corner maxima, the MinE has just
come off the upper corner maxima.  It will diffuse into the right
corner and left corner.  In the left corner it interacts with the MinD
bound to the membrane there.  In the right corner it sits and
eventually diffuses away, but at a somewhat slow pace.  \fixme{An idea
  is that once the MinD travels back from the left corner after its
  second maxima there, it is not able tot set up a maxima in the upper
  corner as easily, since there is left over MinE there, which has
  just traveled from the right corner.  The problem with this is that
  the MinE doesn't seem to be maximized in the upper corner when the
  MinD is traveling back up from the left.}  The other idea is simply
that because the minE is traveling all the way from the right corner,
the longer time it takes to get there simply means that it all
concentrates on the left, leaving the upp and right side completely
open for the build up of MinD.  \fixme{Still dont quite get it but
  take a look at the other simulations and come back.} Actually, you
can see this effect very slightly watching the 98 shape, between the
center cap and the right corner.

Simulations of this shape that has the same height and dimensions .6
times the size (cells roughly $1.6\mu m$ from top to bottom) do not
exhibit the same pattern.  In these the MinD maxima oscillate quickly
back and forth between the two lower corner poles.

Simulations in which the two planar dimensions are 2 and then 3 times
the size (cells roughly $6\micron$ and $9\micron$, respectively, from top
to bottom) show no stable regular periodic pattern \fixme{although
  they haven't really run long enough to exhibit much.  They certainly
  won't show the double oscillation pattern, however.}  The animations
of $6\micron$ cells show that the MinD acts as a sort of traveling
cluster that is constantly being eaten away be MinE behind it.  When
this cluster enters a region of high curvature that ends in a dead end
corner it performs the polar maxima behavoir seen in the other
simulations, regrouping out away from the corner where it travels in
the opposite direction while being ``chased'' away by the MinE that
has itself maximized in the corner.  If it has just maximized in the
top corner, it regroups traveling downward and then naturally splits
into two groups that travel out and form maxima in the two opposite
bottom corners.  After maximizing in these two corners the two groups
recombine again in the center and travel in some direction together as
one cluster.  Simulations of the $9\micron$ cells show a similar roaming
cluster.  Although here, the maxima occur on non-polar walls often,
and will often split into two roaming clusters that interact less
quickly and are less correlated than those seen in the $6\micron$ cell.

It seems that there is an approximate critical cell size for which the
MinD will act in these ``wandering cluster'' formations.  The cell is
large enough so that there is a loss of MinD correlation between the
opposite sides (the protein in the right corner at any time is too far
away from the left to immediately affect the protein there).

%% \begin{center}
%%   \begin{figure}
%%     {\includegraphics[width=\paperwidth]{../data/shape-randst/plots/image-plot_E--randst-25-602-602-9900-1500}}
%%     \caption{Attempt at adding image plot to paper.}
%%     \label{image-plot}
%%   \end{figure}
%% \end{center}

\subsubsection{Randst-97-}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-800-600-9700-1500}
  %%\includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9600-1500}
  \caption{Total protein fluctuation in the sideways flying saucer
    pancake cell shape.  The vertical axis shows stacked the total
    number of proteins that are of four different compound stages and
    in the different sections of the cell.}
  \label{box-97}
\end{figure}

Early on in its simulation the sideways flying saucer pancake shape
shows a clear breakdown of an unstable equilibrium oscillation.  We
have set the intial conditions so that MinD density is highly
concentrated to the left of a dividing line that runs parrallel to the
left vertical side.  For the first number of oscillations the maxima
trade off right to left, from a high peak in the rightward middle
compartment to vertically opposed peaks in the poles of the long axis
of the cell.  This is seen in the first part of Figure~\ref{box-97},
where the lower and upper sections peak in unison, and the middle
right section peaks $\pi$ radians removed from these.  After a few
oscillations the back and forth oscillations tip toward the vertical
direction and the system loses its initial oscillation pattern.  It
falls quickly instead into a back and forth pattern between the two
opposing endcaps.  The oscillation at this point is reminiscent of the
pills shape oscillation, once again showing the robust nature of the
back and forth, two pole pattern.

During each period, directly after one of the polar maxima has
subsided, there is a small spike with the regular maxima pattern (a
spike in membrane bound MinD:ATP immediately followed by a spike in
membrane bound MinD:MinE:ATP) seen in the small middle right
compartment section.  The animations show that the MinD, oscillating
back and forth between the upper and lower poles, seems to catch and
lag in the indentation for a bit before moving on.  Once again the
positive radius of curvature acts as an attractive place for MinD to
cluster on the walls, and it is not released to move on through the
cytoplasm again until MinE has formed its ring and completely reacted
it away.

\subsubsection{Randst-96-Star}
\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-700-700-9600-1500}
  %%\includegraphics[width=\columnwidth]{../data/shape-randst/plots/box-plot_D--randst-25-600-600-9600-1500}
  \caption{Total protein fluctuation in the star cell shape.  The vertical axis shows stacked the total
    number of proteins that are of four different compound stages and
    in the different sections of the cell.}
  \label{box-96}
\end{figure}

This clover shape finds a regular oscillatory behavoir after about
seven periods that consists of two maxima that peak regularly in the
corners accross from each other at the same times (Figure
\ref{box-96}) and then again opposing maxima that peak at the same
time in the other two corners.\fixme{should try some different
  starting conditions with this to see if there's another patter.}

Looking at at few periods in detail.  We find that there is a small
increase in cytoplasmic MinE always directly before the the beggining
of climb in the membrane bound MinE peak.  Also, there is at any one
time a very large amount more of minE that is membrane bound then
cytoplasmic in any ones section.  A difference of a factor of roughly
20 when membrane MinE is at its peak, and roughly 12 when membrane
MinE is at it's minimum.  There seem to be two modes of oscillation,
each exhibiting a density peak maxima that moves back and forth
diagonaly from one corner to the opposite, as if there were two
overlapping, criss-crossed cells oscillating.  In the box plot this is
seen as roughly coinciding membrane minD peaks in the lower left and
right corners, followed one pi later in peaks in the upper right and
upper left corners.  Interesting that this did not exhibit
counterclockwise motion that Huangs 2008 paper might suggest.  Should
run sims that start in the wacky way that just started running with
triangles.
\newline

We haven't simulated long enough yet, so no real pattern has shown
itself.  There's certainly oscillations, and it's almost as if there's
competing modes - top to bottom, left to right, corner to corner, etc.
Interesting.

\fixme{maybe?}The animations seem to relax into a pattern in which the oscillations
go from the smallest corner, out to the other three, and then back to
the smallest corner?  \fixme{There isn't enough data to see properly
  though, need to run for longer.}

\section{Interpretation of Data}
\section{Conclusion}
\bibliography{paper}

\end{document}

\section*{Appendix}

\section{Below are NOTES for the Writing}

Figure \ref{total-oscillation-triangle-plot} shows the equivalent
motion and transference between compound states for the triangular
states as Figure \ref{total-oscillation-plot} does for the pill shaped
cells.  The equilateral triangular shows that in all three sections
there is a spike in wall attached MinD-ATP directly followed by a
spike in wall attached MinD-MinE-ATP.  The pill had shown a similar
pattern except that the mid section exhibited no such pronounced
spike.  Interestingly, the right and mid sections show spikes at
similar times when compared to the left section spikes, which seem to
oscillate in a manner shifted pi off from the other two.  This
behavoir shows itself in the animations, which show a relaxation into
a stable oscillation from a maxima in one corner to a double maxima in
the other two corners, and back again.  The left section heights are
roughly twice that of the right section heights.  The simulations are
started with higher concentrations of proteins in primarily the right
section, with some overlap into the middle section, so that perhpas
this oscillation pattern seems reasonable.  However, it's interesting
to note that the oscillations, started with assymmetric density, do
not naturally rotate around the corners of the triangle in a circular
pattern.

It seems as well that there is a desired oscillation maxima for the
particular shape and size of the cell, as can be seen by the
relaxation, over the first four periods, to a regular maximal height
of the right and left sections' total protein at the height of each spike.

\fixme{Note to make sure I'm interpreting the iscosolese triangle data
  correctly: the triangle sides are 5,5,3, and the higher density
  starts in the corner that is opposite the short side.  This corner
  is covered by the right section.  I've verified this by looking at
  the protein\_microscopy and drawing a little picture, which is on my
  desk...if you dare look for something on my desk!!!}  The iscosolese
triangle starts its simulation with a higher concentration of protein
in the section that is opposite the shorter side.  It quickly relaxes
into a stable oscillation pattern.  This pattern shows that right
corner section regularly has higher spikes in total protein than the
other sections, and that its spikes appear one pi removed from the
spikes for the other two sections.  This a similar behavoir to that
which is seen in the equilateral triangle, differening in that it
converges more quickly towards this behavoir and that it shows a
greater exageration in the height of protein buildup in the right most
section.

\fixme{Note on interpretation of 3-4-5 triangle: the higher density
  starts to the right of a vertical line that is parrallel to the
  second longest side in the triangle (the longer of the two
  non-hypotonuse sides).  This means the higher concentration covers
  the right angle and also the smaller angle in the triangle} The
3-4-5 triangle exhibits behavoir that I need more data to properly
observe.  I does not converge as quickly as that for the iscosolese or
equilateral triangles.

\end{document}